# Simulation Study: Bayesian vMF models for current direction Data with p=2

This vignette demonstrates a complete simulation study using `bayesdir` on synthetic traffic data from a grid road network.

```{r setup, include = FALSE,eval = FALSE}
library(bayesdir)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork) 
library(RColorBrewer)
library(rstiefel) 
library(circular) ## for the coordinates in polar system 
library(circglmbayes) ## for sampling kappa through mcmc method. 
library(parallel)
library(coda) ## for effienct sample size

set.seed(20260509)
```

------------------------------------------------------------------------

## Sythetic data for illustration with p=2

Generate data from von-mises fisher distribution. return in both cartesian coordinate and polar coordinate.

Here $\mu \in [0,2\pi]$ as an angle. $\kappa \geq0$ as concentration.

```{r,eval=FALSE,include=FALSE,purl=FALSE}
#' function to generate sythetic data with p=2
#' @return 
generate_vmf_data <- function(n, mu_true, kappa_true) {
  
  eta_true <- kappa_true * c(cos(mu_true), sin(mu_true))
  raw_samples <- replicate(n, rstiefel::rmf.vector(eta_true))[,1,]
  X_cartesian <- matrix(raw_samples, nrow = 2, ncol = n)
  theta_rad <- atan2(X_cartesian[2, ], X_cartesian[1, ])
  list(X = X_cartesian, theta = theta_rad)
}
```

An example,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
data <- generate_vmf_data(100, 2, 4)
```

For p \>2, simply follow $\texttt{rstiefel::rmf.vector}$, below we have an example.

```{r,eval=FALSE,include=FALSE,purl=FALSEr}
X <- replicate(n,rmf.vector(c(1,1,1)))
```

## Bayesian Inference on a conjugate prior with p=2.

Used as an benchmark. See Forbes & Mardia (2015) by $\texttt{circglmbayes}$.

For the von Mises distribution on $[0,2\pi)$, $f(\theta\mid\mu,\kappa) = \frac{1}{2\pi I_0(\kappa)}\exp\{\kappa\cos(\theta-\mu)\}$, and conjugate prior $\pi(\mu,\kappa) \propto I_0(\kappa)^{-c}\exp\{\kappa R_0\cos(\mu-\mu_0)\},$ with hyperparameters $c\ge0$, $R_0\ge0$, and $\mu_0$. The resulting joint posterior is$$\pi(\mu,\kappa\mid\theta_{1:n}) \propto I_0(\kappa)^{-(c+n)}\exp{\kappa R_n\cos(\mu-\bar\theta_n)}$$ where the posterior resultant length $R_n$ and mean direction $\bar\theta_n$ are defined by $R_n\cos\bar\theta_n = R_0\cos\mu_0 + \sum_{i=1}^{n}\cos\theta_i$ and $R_n\sin\bar\theta_n = R_0\sin\mu_0 + \sum_{i=1}^{n}\sin\theta_i$.

We sample from the posterior using a two‑step Gibbs sampler.\
Given current values $(\mu,\kappa)$,

- Draw $\mu\mid\kappa,\theta$ from a von Mises distribution with mean direction $\bar\theta_n$ and concentration $\kappa R_n$.\

- Draw $\kappa\mid\mu,\theta$ from the full conditional $$    \pi(\kappa\mid\mu,\theta) \propto I_0(\kappa)^{-\eta}\exp\{-\eta\beta_0\kappa\},
    $$ where $\eta = c+n$ and $\beta_0 = -\frac{R_n}{\eta}\cos(\mu-\bar\theta_n)$.

Here $\mu \in [0,2\pi]$ as an angle. $\kappa \geq0$ as concentration. Also, $\texttt{theta}$ as an vector with each scalar $\theta_i\in [0, 2\pi]$ and dimension n.

```{r,eval=FALSE,include=FALSE,purl=FALSE}
#' general MCMC fitter
#' @param theta data
#' @param prior list with hyperparameter (assume to be non informative)
#' @param n_iter,burn number of iteration and burn-in, (n_iter > burn)
#' @return a (n_iter-burn)x2 matrix where returns (n_iter-burn)  samples.
fit_mcmc_vmf <- function(theta, prior = list(c = 0, R0 = 0, mu0 = 0), n_iter = 10000, burn = 5000) {
  n <- length(theta)
  C_n <- prior$R0 * cos(prior$mu0) + sum(cos(theta))
  S_n <- prior$R0 * sin(prior$mu0) + sum(sin(theta))
  Rn <- sqrt(C_n^2 + S_n^2)
  theta_bar_n <- atan2(S_n, C_n)
  eta_param <- prior$c + n
  
  mu_sam <- numeric(n_iter); kappa_sam <- numeric(n_iter)
  mu_sam[1] <- theta_bar_n; kappa_sam[1] <- 1
  
  for (iter in 2:n_iter) {
    mu_sam[iter] <- as.numeric(circular::rvonmises(1, circular::circular(theta_bar_n), kappa_sam[iter - 1] * Rn))
    etag <- -Rn * cos(mu_sam[iter] - theta_bar_n)
    kappa_sam[iter] <- circglmbayes:::sampleKappa(etag = etag, eta = eta_param)[1]
  }
  cbind(mu = mu_sam[-(1:burn)], kappa = kappa_sam[-(1:burn)])
}
```

Below we will show an brief example.

```{r,eval=FALSE,include=FALSE,purl=FALSE}
mu_true <- 2
kappa_true <- 4
data <- generate_vmf_data(100, mu_true = mu_true, kappa_true = kappa_true)
res_MCMC <- fit_mcmc_vmf(data$theta)

kappa_MCMC <- res_MCMC[,2]
mu_MCMC <- res_MCMC[,1]

ess_mu    <- effectiveSize(mu_MCMC)
ess_kappa <- effectiveSize(kappa_MCMC)
cat("Effective Sample Size (after burn-in):\n")
cat(sprintf("  mu    : %.1f\n", ess_mu))
cat(sprintf("  kappa : %.1f\n", ess_kappa))
```

Visualize Result after fitting data,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
plot(kappa_MCMC,type = "l",xlab = "Iteration", ylab = "kappa", main = "Trace of kappa")
abline(h = kappa_true,lty=4,col = "red")

kappa_den <- density(kappa_MCMC)
plot(kappa_den,main = "Posterior density of kappa", xlab = "kappa",
         col = "steelblue", lwd = 2)
abline(v = kappa_true,lty=4,col = "red")


plot(mu_MCMC,type = 'l',xlab = "Iteration", ylab = "mu", main = "Trace of mu")
abline(h = mu_true,lty=4,col = "red")

mu_den <- density(mu_MCMC)
plot(mu_den,main = "Posterior density of mu", xlab = "mu",
         col = "steelblue", lwd = 2)
abline(v = mu_true,lty=4,col = "red")
```

## A simple example for MPS to help you draw graphies in the paper

Though $\texttt{bayesdir::run_predresamp_vmf()}$ in $\texttt{bayesdir()}$

```{r,eval=FALSE,include=FALSE,purl=FALSE}
X <- data$X
nn_cores <- max(detectCores()-1,1)
res_mart <- bayesdir::run_predresamp_vmf(X,B=500,M=1000,n_cores = nn_cores)

kappa_mart <- res_mart$sample$kappa_sam
mu_mart <- res_mart$sample$mu_sam
```

Visualize Result after fitting data,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
kappa_den <- density(kappa_mart)
plot(kappa_den,main = "Posterior density of kappa", xlab = "kappa",
         col = "steelblue", lwd = 2)
abline(v = kappa_true,lty=4,col = "red")


mu_den <- density(mu_mart)
plot(mu_den,main = "Posterior density of mu", xlab = "mu",
         col = "steelblue", lwd = 2)
abline(v = mu_true,lty=4,col = "red")
```

For simplicity, we only perform correction on last point in $\texttt{bayesdir::run_predresamp_vmf()}$. and $\texttt{bayesdir::run_single_martingale()}$. So to draw similar graphes with the paper, a modified functions are needed, as $\texttt{run_cor_mart()}$, $\texttt{run_cor_predresamp_vmf()}$.(If you do not want to recap this part, just ignore it.)

```{r,eval=FALSE,include=FALSE,purl=FALSE}
#' Run a single martingale posterior chain (no preconditioner).
#' @param c0 initial natural parameter vector (use MLE)
#' @param N number of observed data points
#' @param M number of iteration
#' @param save_path logical, if TRUE the corrected chain is also returned
run_cor_martingale <- function(c0, N, M = 1000, save_path = TRUE) {
  p <- length(c0)
  c_curr <- c0            # working chain (uncorrected)
  V_mat <- matrix(0, p, p)
  total_w <- 0

  if (save_path) {
    x_chain <- matrix(0, p, M)
    c_chain <- matrix(0, p, M)   # will store corrected values
  }

  # function for stable inversion
  safe_sym_inverse <- function(M, tol = 1e-10) {
    eig <- eigen((M + t(M)) / 2, symmetric = TRUE)
    vals_inv <- 1 / pmax(eig$values, tol)
    eig$vectors %*% diag(vals_inv) %*% t(eig$vectors)
  }

  # main loop
  for (t in 1:M) {
    # generate new observation
    x_new <- rstiefel::rmf.vector(c_curr)
    s <- score_fn(c_curr, x_new, p)
    wt <- 1 / (t + N - 1)
    total_w <- total_w + wt^2
    V_mat <- V_mat + wt^2 * get_FIM_vmf(c_curr, p)
    # update the working parameter (no correction)
    c_curr <- c_curr + wt * s

    if (save_path) {
      x_chain[, t] <- x_new
      # correct the current c_curr using accumulated information
      A_t <- total_w * safe_sym_inverse(V_mat, tol = 1e-10)
      c_curr_corrected <- c0 + A_t %*% (c_curr - c0)
      c_chain[, t] <- c_curr_corrected
    }
  }

  # final correction (same as before)
  delta <- c_curr - c0
  A <- total_w * safe_sym_inverse(V_mat, tol = 1e-10)
  c_final <- c0 + A %*% delta

  # convert to polar coordinates (for p = 2)
  kappa_final <- sqrt(sum(c_final^2))
  if (p == 2) {
    mu_final <- atan2(c_final[2], c_final[1])
    if (mu_final < 0) mu_final <- mu_final + 2 * pi
  } else {
    mu_final <- (1 / kappa_final) * c_final
  }

  res <- list(
    c_final     = c_final,
    mu_final    = mu_final,
    kappa_final = kappa_final
  )
  if (save_path) {
    res$x_chain <- x_chain
    res$c_chain <- c_chain
  }
  return(res)
}

run_cor_predresamp_vmf <- function(X,B = 5000, M = 1000, n_cores = 1,return_chains = TRUE){
  
  n  <- ncol(X)
  p  <- nrow(X)
  c0 <- mle_vmf(X)
  
  
  chains <- parallel::mclapply(1:B, function(b) {
    run_cor_martingale(c0 = c0, N = n, M = M, save_path = return_chains)
  }, mc.cores = n_cores)
  
  c_sam <- sapply(chains, `[[`, "c_final")
  kappa_sam <-sapply(chains,`[[`, "kappa_final")
  mu_sam <-sapply(chains,`[[`, "mu_final")
  
  sample_list <- list(
    c_sam     = c_sam,
    mu_sam    = mu_sam,
    kappa_sam = kappa_sam
  )

  out <- list(sample = sample_list)
  if (return_chains) {
    out$chain <- chains
  }
  
  out
}
```

plot function

```{r}
plot_trace_sideways_density <- function(
    paths,
    samples,
    iter = NULL,
    ref_value = NULL,
    ref_density_fun = NULL,
    xlab = "Forward step i",
    ylab = expression(hat(mu)[n+i]),
    main = NULL,
    trace_col = rgb(0.25, 0.45, 0.75, 0.06),
    trace_lwd = 0.5,
    density_fill = rgb(0.25, 0.45, 0.75, 0.30),
    density_border = rgb(0.20, 0.35, 0.65),
    density_lwd = 1.2,
    ref_col = "red",
    ref_lwd = 1.5,
    ref_lty = 2,
    ref_density_col = "red",
    ref_density_lwd = 2,
    density_width_frac = 0.22,
    density_x_offset_frac = 1.02,
    xlim_multiplier = 1.25,
    y_pad_frac = 0.04,
    cex_lab = 1.2,
    cex_axis = 1.0,
    cex_main = 1.2,
    cex_leg = 0.9,
    legend_pos = "topleft",
    show_legend = TRUE,
    show_traj_legend = FALSE,
    mark_initial = TRUE,
    initial_col = "red",
    initial_pch = 19,
    initial_cex = 1.2,
    initial_label = "initial point",
    n_value = 500,
    n_pos = "bottomright",
    n_cex = 0.8,
    n_col = "gray40",
    n_pad_x = 0.02,   
    n_pad_y = 0.03    
) {
  if (!is.matrix(paths)) paths <- as.matrix(paths)
  if (!is.numeric(paths)) stop("`paths` must be numeric.")
  if (!is.numeric(samples)) stop("`samples` must be numeric.")

  if (is.null(iter)) iter <- seq_len(ncol(paths)) - 1
  if (length(iter) != ncol(paths)) {
    stop("`iter` must have length equal to `ncol(paths)`.")
  }

  x_start <- min(iter)
  x_end <- max(iter)
  x_width <- x_end - x_start
  if (x_width <= 0) stop("`iter` must span a positive range.")

  y_range <- range(c(as.vector(paths), samples, ref_value), finite = TRUE)
  y_pad <- y_pad_frac * diff(y_range)
  if (!is.finite(y_pad) || y_pad == 0) y_pad <- 0.1
  y_range <- y_range + c(-y_pad, y_pad)

  dens <- density(samples, n = 512)
  y_grid <- seq(y_range[1], y_range[2], length.out = 500)

  ref_density <- NULL
  if (!is.null(ref_density_fun)) ref_density <- ref_density_fun(y_grid)

  dens_scale <- x_width * density_width_frac / max(dens$y, ref_density, na.rm = TRUE)
  x_offset <- x_start + x_width * density_x_offset_frac
  x_max <- x_start + x_width * xlim_multiplier

  plot(
    NULL,
    xlim = c(x_start, x_max),
    ylim = y_range,
    xlab = "",
    ylab = ylab,
    main = main,
    xaxt = "n",
    cex.lab = cex_lab,
    cex.axis = cex_axis,
    cex.main = cex_main
  )

  ticks <- pretty(c(x_start, x_end))
  ticks <- ticks[ticks >= x_start & ticks <= x_end]
  axis(1, at = ticks, labels = ticks, lwd = 0, lwd.ticks = 1, cex.axis = cex_axis)
  usr <- par("usr")
  segments(x_start, usr[3], x_end, usr[3], xpd = FALSE)
  mtext(xlab, side = 1, line = 2.8, at = (x_start + x_end) / 2, cex = cex_lab)

  for (i in seq_len(nrow(paths))) {
    lines(iter, paths[i, ], col = trace_col, lwd = trace_lwd)
  }

  if (!is.null(ref_value)) {
    abline(h = ref_value, col = ref_col, lwd = ref_lwd, lty = ref_lty)
  }

  polygon(
    x_offset + dens$y * dens_scale,
    dens$x,
    col = density_fill,
    border = density_border,
    lwd = density_lwd
  )

  if (!is.null(ref_density)) {
    lines(
      x_offset + ref_density * dens_scale,
      y_grid,
      col = ref_density_col,
      lwd = ref_density_lwd
    )
  }

  segments(x_offset, y_range[1], x_offset, y_range[2],
           col = "gray60", lty = 3, lwd = 0.8)

  if (mark_initial) {
    init_x <- iter[1]
    init_y <- paths[1, 1]
    points(init_x, init_y, col = initial_col, pch = initial_pch, cex = initial_cex)
  }

  
  if (!is.null(n_value)) {
    usr <- par("usr")   # c(xmin, xmax, ymin, ymax)
    x_range <- usr[2] - usr[1]
    y_range_cur <- usr[4] - usr[3]

    if (n_pos == "bottomright") {
      x_pos <- usr[2] - x_range * n_pad_x   
      y_pos <- usr[3] + y_range_cur * n_pad_y   
      adj <- c(1, 0)
    } else if (n_pos == "bottomleft") {
      x_pos <- usr[1] + x_range * n_pad_x
      y_pos <- usr[3] + y_range_cur * n_pad_y
      adj <- c(0, 0)
    } else {
      warning("n_pos should be 'bottomright' or 'bottomleft'. Defaulting to bottomright.")
      x_pos <- usr[2] - x_range * n_pad_x
      y_pos <- usr[3] + y_range_cur * n_pad_y
      adj <- c(1, 0)
    }

    text(x = x_pos, y = y_pos, labels = paste0("n = ", n_value),
         adj = adj, cex = n_cex, col = n_col)
  }

  if (show_legend) {
    leg_text <- c("True value", "MPS density", "MCMC density")
    leg_col  <- c(ref_col, density_border, ref_density_col)
    leg_lty  <- c(ref_lty, 1, 1)
    leg_lwd  <- c(ref_lwd, density_lwd, ref_density_lwd)
    leg_pch  <- c(NA, 15, NA)
    leg_pt.cex <- c(NA, 1.5, NA)

    if (show_traj_legend) {
      leg_text <- c("MPS trajectories", leg_text)
      leg_col  <- c(adjustcolor(trace_col, alpha.f = 0.6), leg_col)
      leg_lty  <- c(1, leg_lty)
      leg_lwd  <- c(1.5, leg_lwd)
      leg_pch  <- c(NA, leg_pch)
      leg_pt.cex <- c(NA, leg_pt.cex)
    }

    if (mark_initial) {
      leg_text <- c(leg_text, initial_label)
      leg_col  <- c(leg_col, initial_col)
      leg_lty  <- c(leg_lty, NA)
      leg_lwd  <- c(leg_lwd, NA)
      leg_pch  <- c(leg_pch, initial_pch)
      leg_pt.cex <- c(leg_pt.cex, 1.2)
    }

    legend(
      legend_pos,
      legend = leg_text,
      col    = leg_col,
      lty    = leg_lty,
      lwd    = leg_lwd,
      pch    = leg_pch,
      pt.cex = leg_pt.cex,
      bty    = "n",
      cex    = cex_leg
    )
  }

  invisible(list(
    density = dens,
    y_range = y_range,
    x_offset = x_offset,
    dens_scale = dens_scale
  ))
}
```

Example for drawing graphes.

```{r}

set.seed(2026) #for n=100,500

#Settings
nn_cores <- max(detectCores()-1,1)
mu_true <- 2
kappa_true <- 4
n <- 500 # use n=500 for another graph
data <- generate_vmf_data(n, mu_true = mu_true, kappa_true = kappa_true)

#MCMC
res_MCMC <- fit_mcmc_vmf(data$theta,n_iter=7000,burn = 5000)
kappa_MCMC <- res_MCMC[,2]
kappa_den <- density(kappa_MCMC)
ref_fun <- approxfun(kappa_den$x, kappa_den$y, yleft = 0, yright = 0)

#MPS
cor_res <- run_cor_predresamp_vmf(data$X, B = 2000, M = 1000, n_cores = nn_cores, return_chains = TRUE)
kappa_paths <- t(sapply(cor_res$chain, function(chain) {
  sqrt(colSums(chain$c_chain^2))
}))
c0 <- bayesdir::mle_vmf(data$X)   
kappa0 <- sqrt(sum(c0^2))
kappa_paths_with_init <- cbind(kappa0, kappa_paths)
kappa_last <- kappa_paths_with_init[, ncol(kappa_paths_with_init)]

#Draw graph, Finally!
plot_trace_sideways_density(
  paths          = kappa_paths_with_init,
  samples        = kappa_last,          
  iter           = 1:1001,
  ref_value      = kappa_true,          
  ref_density_fun = ref_fun,            
  ref_density_col = "blue",             
  ref_density_lwd = 2,                  
  xlab = "Forward step",
  ylab = "kappa",
  cex_leg = 0.7,
  n_value = n,
  n_cex = 0.9
)

```

## Comparison between MCMC and MPS Method.

As usual, here first give out the function.

```{r,eval=FALSE,include=FALSE,purl=FALSE}
#' Compare MCMC method with Bayesdir method.
#' @param n number of observations
#' @param n_sim number of simulation
#' @param mu_true,kappa_true ground truth
#' @param B number of samples from each sampler
#' @param M length of each martingale in MPS
#' @param prior prior in MCMC method
#' @param n_cores number of cores you can occupy
#' @param save_dir,save_interval out of safety, save result in xxx each xxx times.
sim_compare_vmf <- function(n, n_sim = 1000,
                            mu_true = 2, kappa_true = 4,
                            B = 5000, M = 1000,
                            prior = list(c = 0, R0 = 0, mu0 = 0),
                            seed = 1234, n_cores = 1,
                            save_dir = "sim_results",
                            save_interval = n_sim/5) {
  
  
  if (!dir.exists(save_dir)) {
    dir.create(save_dir, recursive = TRUE)
  }
  
  set.seed(seed)
  
  
  get_stats <- function(samples, true_val) {
    ci <- quantile(samples, probs = c(0.025, 0.975))
    list(
      cover = as.numeric(true_val >= ci[1] && true_val <= ci[2]),
      len = as.numeric(ci[2] - ci[1]),
      est = var(samples) 
    )
  }
  
  pb <- txtProgressBar(min = 0, max = n_sim, style = 3, 
                       label = "Simulation Progress: ")
  
  start_time <- Sys.time()
  res_mart <- vector("list", n_sim)
  res_bayes <- vector("list", n_sim)

  for (i in 1:n_sim) {
    
    # Generate Data
    data <- generate_vmf_data(n, mu_true, kappa_true)
    
    # --- Martingale Posterior ---
    t0 <- proc.time()[3]
    m_chains <- run_predresamp_vmf(data$X, B = B, M = M, n_cores = n_cores)
    t1 <- proc.time()[3]
    
    m_mu <- get_stats(m_chains$sample$mu_sam, mu_true)
    m_ka <- get_stats(m_chains$sample$kappa_sam, kappa_true)
    
    res_mart[[i]] <- data.frame(
      c_mu = m_mu$cover, l_mu = m_mu$len, e_mu = m_mu$est,
      c_ka = m_ka$cover, l_ka = m_ka$len, e_ka = m_ka$est,
      time = t1 - t0
    )

    # --- Bayes MCMC ---
    t2 <- proc.time()[3]
    b_samples <- fit_mcmc_vmf(data$theta, prior, n_iter = B + 5000, burn = 5000)
    t3 <- proc.time()[3]
    
    b_mu <- get_stats(b_samples[, "mu"], mu_true)
    b_ka <- get_stats(b_samples[, "kappa"], kappa_true)
    
    res_bayes[[i]] <- data.frame(
      c_mu = b_mu$cover, l_mu = b_mu$len, e_mu = b_mu$est,
      c_ka = b_ka$cover, l_ka = b_ka$len, e_ka = b_ka$est,
      time = t3 - t2
    )
    
    setTxtProgressBar(pb, i)

    
    if (i %% save_interval == 0) {
      temp_res <- list(
        mart = res_mart[1:i], 
        bayes = res_bayes[1:i],
        current_index = i,
        params = list(n=n, mu_true=mu_true, kappa_true=kappa_true),
        save_time = Sys.time()
      )
      file_name <- file.path(save_dir, paste0("backup_iter_", i, ".rds"))
      saveRDS(temp_res, file = file_name)
      message(sprintf("\n[Checkpoint] Saved at %s | Elapsed: %.1f mins | ETA: %.1f mins (Finish approx. %s)", 
                format(Sys.time(), "%H:%M:%S"), 
                as.numeric(difftime(Sys.time(), start_time, units = "mins")), 
                (as.numeric(difftime(Sys.time(), start_time, units = "mins")) / i) * (n_sim - i), 
                format(Sys.time() + ((as.numeric(difftime(Sys.time(), start_time, units = "secs")) / i) * (n_sim - i)), "%H:%M")))
    }
  }
  
  close(pb)

  df_mart <- do.call(rbind, res_mart)
  df_bayes <- do.call(rbind, res_bayes)

  summarize_res <- function(df) {
    data.frame(
      n = n,
      Cov_mu  = mean(df$c_mu) * 100,
      Len_mu  = mean(df$l_mu),
      Max_Var_mu  = max(df$e_mu),  
      Cov_ka  = mean(df$c_ka) * 100,
      Len_ka  = mean(df$l_ka),
      Max_Var_ka  = max(df$e_ka), 
      Mean_Time = mean(df$time)
    )
  }

  final_summary <- rbind(
    cbind(Method = "Martingale", summarize_res(df_mart)),
    cbind(Method = "BayesMCMC", summarize_res(df_bayes))
  )

  return(final_summary)
}
```

As an simple example,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
getwd()
setwd(".") ## here set folder to save result, also it will be the place holder to save intermediate result for safety.
nn_cores <- max(detectCores()-1,1)
res <- sim_compare_vmf(n=100,n_sim=500,B=500,M=1000,n_cores = nn_cores)
```

As for code in the paper,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
getwd()
setwd(".") ## here set folder to save result, also it will be the place holder to save intermediate result for safety.
nn_cores <- max(detectCores()-1,1)
res500 <- sim_compare_vmf(n=500,n_sim=5000,B=5000,M=1000,n_cores = nn_cores)
save(res500, file = "vmf_simulation_results.RData")

res200 <- sim_compare_vmf(n=200,n_sim=5000,B=5000,M=1000,n_cores = nn_cores)
save(res500, file = "vmf_simulation_results.RData")

res100 <- sim_compare_vmf(n=100,n_sim=5000,B=5000,M=1000,n_cores = nn_cores)
save(res500, file = "vmf_simulation_results.RData")

res10 <- sim_compare_vmf(n=10,n_sim=5000,B=5000,M=1000,n_cores = nn_cores)
save(res500, file = "vmf_simulation_results.RData")
```
