# Simulation Study: Bayesian vMF models for current direction Data

This vignette demonstrates a complete simulation study using `bayesdir` on synthetic traffic data from a grid road network.

```{r setup, include = FALSE,eval = FALSE}
library(bayesdir)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
library(RColorBrewer)
library(rstiefel)
library(circular)
library(circglmbayes)
library(parallel)

set.seed(20260509)
```

------------------------------------------------------------------------

## Sythetic data for illustration with p=2

------------------------------------------------------------------------

## Bayesian Inference on a conjugate prior.

Used as an benchmark. See Forbes & Mardia (2015) by $\texttt{circglmbayes}$

Here $\mu \in [0,2\pi]$ as an angle. $\kappa \geq0$ as concentration.

```{r,eval=FALSE,include=FALSE,purl=FALSE}
sim_bayes_vm <- function(n, n_sim = 100,
                         mu_true = 2, kappa_true = 4,
                         prior = list(c = 0, R0 = 0, mu0 = 0),
                         n_iter = 10000, burn = 5000,
                         save_chains = FALSE,   
                         seed = 2024) {

  
  c   <- prior$c
  R0  <- prior$R0
  mu0 <- prior$mu0
  
  cover_mu    <- numeric(n_sim)
  cover_kappa <- numeric(n_sim)
  len_mu      <- numeric(n_sim)
  len_kappa   <- numeric(n_sim)
  mean_mu     <- numeric(n_sim)
  mean_kappa  <- numeric(n_sim)
  mcse_mu     <- numeric(n_sim)
  mcse_kappa  <- numeric(n_sim)
  runtimes    <- numeric(n_sim)
  
  if (save_chains) {
    chains_list <- vector("list", n_sim)
  }
  
  set.seed(seed)
  
  for (i in 1:n_sim) {
    theta <- as.numeric(rvonmises(n, circular(mu_true), kappa_true))
    

    C_n <- R0 * cos(mu0) + sum(cos(theta))
    S_n <- R0 * sin(mu0) + sum(sin(theta))
    Rn <- sqrt(C_n^2 + S_n^2)
    theta_bar_n <- atan2(S_n, C_n)
    eta <- c + n
    

    Rbar <- Rn / n
    if (Rbar < 0.001) {
      kappa_init <- 0
    } else if (Rbar > 0.999) {
      kappa_init <- 1 / (2 * (1 - Rbar))
    } else {
      solve_kappa <- function(k) besselI(k, 1, expon.scaled = TRUE) / besselI(k, 0, expon.scaled = TRUE) - Rbar
      kappa_init <- uniroot(solve_kappa, c(0.001, 100))$root
    }
    mu_init <- theta_bar_n
    
    mu_sam    <- numeric(n_iter)
    kappa_sam <- numeric(n_iter)
    mu_sam[1]    <- mu_init
    kappa_sam[1] <- kappa_init
    
    rvm_one <- function(mean_rad, kappa) {
      as.numeric(rvonmises(1, circular(mean_rad), kappa))
    }
    
    t0 <- proc.time()[3]
    for (iter in 2:n_iter) {
      mu_sam[iter] <- rvm_one(theta_bar_n, kappa_sam[iter - 1] * Rn)
      etag <- -Rn * cos(mu_sam[iter] - theta_bar_n)
      kappa_sam[iter] <- circglmbayes:::sampleKappa(etag = etag, eta = eta)[1]
    }
    t1 <- proc.time()[3]
    runtimes[i] <- t1 - t0
    

    mu_post    <- mu_sam[-(1:burn)]
    kappa_post <- kappa_sam[-(1:burn)]
    B_post <- length(mu_post)
    
    mean_mu[i]    <- mean(mu_post)
    mean_kappa[i] <- mean(kappa_post)
    
    
    ci_mu    <- quantile(mu_post, c(0.025, 0.975))
    ci_kappa <- quantile(kappa_post, c(0.025, 0.975))
    
    cover_mu[i]    <- (mu_true >= ci_mu[1] && mu_true <= ci_mu[2])
    len_mu[i]      <- ci_mu[2] - ci_mu[1]
    cover_kappa[i] <- (kappa_true >= ci_kappa[1] && kappa_true <= ci_kappa[2])
    len_kappa[i]   <- ci_kappa[2] - ci_kappa[1]
    

    if (save_chains) {
      chains_list[[i]] <- cbind(mu = mu_sam, kappa = kappa_sam)
    }
  }
  
 
  summary_df <- data.frame(
    n = n,
    Coverage_mu    = mean(cover_mu) * 100,
    Length_mu      = mean(len_mu),
    Coverage_kappa = mean(cover_kappa) * 100,
    Length_kappa   = mean(len_kappa),
    SE_Mean_mu    = sd(mean_mu),
    SE_Mean_kappa = sd(mean_kappa),
    Mean_Time       = mean(runtimes)
  )
  
  if (save_chains) {
    return(list(summary = summary_df, chains = chains_list))
  } else {
    return(summary_df)
  }
}
```

```{r, eval=FALSE}

#res_bayes30  <- sim_bayes_vm(n = 30, n_sim = 10)
#res_bayes100 <- sim_bayes_vm(n = 100, n_sim = 10)

#result_table <- rbind(res_bayes10, res_bayes30, res_bayes100)
#print(result_table, digits = 3)
```

## Predictive Posterior via Martingale.

Score based martingale sampling without pre-conditioner (inverse info).

For efficiency, hybrid trick is adopted.

```{r, send_to = 'R/simulation.R'}
#' functions for simulations study
#' @param n sample size
#' @param n_sim sim times
#' @param mu_ture,kappa_true true param
#' @param B number of chains
#' @param M number of iteration for each chain
#' @param seed 
#' @param n_cores number of cores you can use
#' @param save_all save data or not
#' @return all the data with a summary we gonna used
#' @export
sim_martingale_study <- function(n, n_sim = 500,
                                 mu_true = 2, kappa_true = 4,
                                 B = 5000,M = 50,
                                 seed = 1234,
                                 n_cores = 1,
                                 save_all = FALSE) {

  RNGkind("L'Ecuyer-CMRG")
  set.seed(seed)
  
  # ---- Storage ----
  cover_mu <- cover_kappa <- numeric(n_sim)
  len_mu <- len_kappa <- numeric(n_sim)
  mean_mu_vec <- mean_kappa_vec <- numeric(n_sim)
  times <- numeric(n_sim)
  if (save_all) {results <- vector("list", n_sim)}
  
  eta_true <- kappa_true * c(cos(mu_true), sin(mu_true))
  
  
  # ---- Simulation loop ----
  for (i in 1:n_sim) {
    # Generate observations
    X <- replicate(n,rstiefel::rmf.vector(eta_true))
    
    
    # Run B independent chains
    t0 <- proc.time()[3]
    chains<- run_predresamp_vmf(X, B = 5000, M=50,n_cores = n_cores,
                                return_chains = save_all)
    t1 <- proc.time()[3]
    times[i] <- t1 - t0
    
    # Extract samples
    sample_list <- chains$sample
    
    mu_sam <- sample_list$mu_sam
    kappa_sam <- sample_list$kappa_sam
    
    mean_mu_vec[i] <- mean(mu_sam)
    mean_kappa_vec[i] <- mean(kappa_sam)
    
    # 95% credible intervals for mu and kappa
    ci_mu <- quantile(mu_sam, probs = c(0.025, 0.975))
    cover_mu[i] <- (mu_true >= ci_mu[1] && mu_true <= ci_mu[2])
    len_mu[i] <- ci_mu[2] - ci_mu[1]
    
    ci_kappa <- quantile(kappa_sam, c(0.025, 0.975))
    cover_kappa[i] <- (kappa_true >= ci_kappa[1]) & (kappa_true <= ci_kappa[2])
    len_kappa[i] <- ci_kappa[2] - ci_kappa[1]
    
    if (save_all) {results[[i]]<- chains}
    
  }
  
  # ---- Build output table ----
  summarys <- data.frame(
    n = n,
    Coverage_mu = mean(cover_mu) * 100,
    Length_mu = mean(len_mu),
    Coverage_kappa = mean(cover_kappa) * 100,
    Length_kappa = mean(len_kappa),
    Mean_SE_mu      = sd(mean_mu_vec),
    Mean_SE_kappa   = sd(mean_kappa_vec),
    Mean_Time = mean(times)
  )
  
  if (save_all) {
    list(summary = summarys, data_all = results)
  } else {
    summarys
  }
}

```


Simulation with n_sim = 5000, B = 5000, M = 50

```{r,eval = FALSE, include=FALSE,purl=FALSE}
res_bayes_10 <- sim_bayes_vm(n=10,n_sim = 1, save_chains = TRUE)
res_bayes_30 <- sim_bayes_vm(n=30,n_sim = 1,save_chains = TRUE)
res_bayes_100 <- sim_bayes_vm(n=100,n_sim= 1, save_chains = TRUE)


result_table_bayes <- rbind(res_bayes_10$summary, res_bayes_30$summary, res_bayes_100$summary)
print(result_table_bayes)
```

```{r,eval = FALSE, include=FALSE,purl=FALSE}
n_cores <- detectCores()-2
res_mart10 <- sim_martingale_study(n=10,n_sim = 5000, B = 5000,M = 50,n_cores = n_cores, save_all = TRUE)
res_mart30 <- sim_martingale_study(n=30,n_sim = 5000, B = 5000,M = 50,n_cores = n_cores, save_all = TRUE)
res_mart100 <- sim_martingale_study(n=100,n_sim = 5000, B = 5000,M = 50,n_cores = n_cores, save_all = TRUE)

result_table_mart <- rbind(res_mart10$summary, res_mart30$summary, res_mart100$summary)
print(result_table_mart)
```
