# Simulation Study: Bayesian vMF models for current direction Data

This vignette demonstrates a complete simulation study using `bayesdir` on synthetic traffic data from a grid road network.

```{r setup, include = FALSE}
library(bayesdir)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
library(RColorBrewer)
library(circular)
library(circglmbayes)

set.seed(20260509)
```

------------------------------------------------------------------------

## Sythetic data for illustration

------------------------------------------------------------------------

## Bayesian Inference on a conjugate prior.

Used as an benchmark. See Forbes & Mardia (2015) by $\texttt{circglmbayes}$

Here $\mu \in [0,2\pi]$ as an angle. $\kappa \geq0$ as concentration.

```{r}
sim_bayes_vm <- function(n, n_sim = 100,
                         mu_true = 2, kappa_true = 4,
                         prior = list(c = 0, R0 = 0, mu0 = 0),
                         n_iter = 10000, burn = 5000,
                         seed = 2024) {
  
  
  c   <- prior$c
  R0  <- prior$R0
  mu0 <- prior$mu0
  
  cover_mu   <- numeric(n_sim)
  cover_kappa <- numeric(n_sim)
  len_mu     <- numeric(n_sim)
  len_kappa   <- numeric(n_sim)
  mean_mu    <- numeric(n_sim)
  mean_kappa  <- numeric(n_sim)
  runtimes   <- numeric(n_sim)
  
  set.seed(seed)
  
  for (i in 1:n_sim) {
    theta <- as.numeric(rvonmises(n, circular(mu_true), kappa_true))
    
    C_n <- R0*cos(mu0) + sum(cos(theta))
    S_n <- R0*sin(mu0) + sum(sin(theta))
    Rn <- sqrt(C_n^2 + S_n^2)
    theta_bar_n <- atan2(S_n, C_n)
    eta <- c + n        
    
    Rbar <- Rn / n
    if (Rbar < 0.001) {
      kappa_init <- 0
    } else if (Rbar > 0.999) {
      kappa_init <- 1 / (2 * (1 - Rbar))
    } else {
      solve_kappa <- function(k) besselI(k,1,expon.scaled=TRUE)/besselI(k,0,expon.scaled=TRUE)-Rbar
      kappa_init <- uniroot(solve_kappa, c(0.001, 100))$root
    }
    mu_init <- theta_bar_n
    
    mu_sam    <- numeric(n_iter)
    kappa_sam <- numeric(n_iter)
    mu_sam[1]    <- mu_init
    kappa_sam[1] <- kappa_init
    
    rvm_one <- function(mean_rad, kappa) {
      as.numeric(rvonmises(1, circular(mean_rad), kappa))
    }
    
    t0 <- proc.time()[3]
    for (iter in 2:n_iter) {
      mu_sam[iter] <- rvm_one(theta_bar_n, kappa_sam[iter-1] * Rn)

      etag <- -Rn * cos(mu_sam[iter] - theta_bar_n)
      kappa_sam[iter] <- circglmbayes:::sampleKappa(etag = etag, eta = eta)[1]
    }
    t1 <- proc.time()[3]
    runtimes[i] <- t1 - t0
    
    mu_post   <- mu_sam[-(1:burn)]
    kappa_post <- kappa_sam[-(1:burn)]
    
    ci_mu <- quantile(mu_post, c(0.025, 0.975))
    ci_kappa <- quantile(kappa_post, c(0.025, 0.975))
    cover_mu[i]   <- (mu_true >= ci_mu[1] & mu_true <= ci_mu[2])
    cover_kappa[i] <- (kappa_true >= ci_kappa[1] & kappa_true <= ci_kappa[2])
    len_mu[i]     <- ci_mu[2] - ci_mu[1]
    len_kappa[i]   <- ci_kappa[2] - ci_kappa[1]
    
    mean_mu[i]   <- mean(mu_post)
    mean_kappa[i] <- mean(kappa_post)
  }
  

  data.frame(
    n = n,
    Parameter = c("mu", "kappa"),
    Coverage  = 100 * c(mean(cover_mu), mean(cover_kappa)),
    Length    = c(mean(len_mu), mean(len_kappa)),
    Max_SE    = c(sd(mean_mu), sd(mean_kappa)),
    Runtime   = mean(runtimes)   
  )
}
```

```{r}
res_bayes10  <- sim_bayes_vm(n = 10, n_sim = 100)
res_bayes30  <- sim_bayes_vm(n = 30, n_sim = 100)
res_bayes100 <- sim_bayes_vm(n = 100, n_sim = 100)

result_table <- rbind(res_bayes10, res_bayes30, res_bayes100)
print(result_table, digits = 3)
```

## Predictive Posterior via Martingale.

Score based martingale sampling without pre-conditioner (inverse info).

For efficiency, hybrid trick is adopted.

```{r}
sim_martingale <- function(n, n_sim = 100,
                                         mu_true = 2, kappa_true = 4,
                                         B = 500, M = 50,
                                         seed = 1234) {
  
  A <- function(k) ifelse(k < 1e-6, 0,
                          besselI(k, 1, expon.scaled = TRUE) /
                            besselI(k, 0, expon.scaled = TRUE))
  
  score_fn <- function(eta, x) {
    k <- sqrt(sum(eta^2))
    if (k < 1e-6) return(x)
    expected <- A(k) * eta / k
    x - expected
  }
  
  rvMF2 <- function(eta) {
    k <- sqrt(sum(eta^2))
    if (k < 1e-6) {
      angle <- runif(1, 0, 2*pi)
    } else {
      mu_angle <- atan2(eta[2], eta[1])
      angle <- as.numeric(rvonmises(1, circular(mu_angle), k))
    }
    c(cos(angle), sin(angle))
  }
  
  Z_list <- list()
  Tn_vec <- c()
  times  <- numeric(n_sim)
  
  set.seed(seed)
  
  for (i in 1:n_sim) {
    theta_data <- as.numeric(rvonmises(n, circular(mu_true), kappa_true))
    
    C <- sum(cos(theta_data)); S <- sum(sin(theta_data))
    R <- sqrt(C^2 + S^2); Rbar <- R/n
    mu_hat <- atan2(S, C)
    if (Rbar < 0.001) kappa_hat <- 0
    else if (Rbar > 0.999) kappa_hat <- 1/(2*(1-Rbar))
    else kappa_hat <- uniroot(function(k) A(k) - Rbar, c(0.001, 100))$root
    eta0 <- kappa_hat * c(cos(mu_hat), sin(mu_hat))
    
    Z_mat <- matrix(0, B, 2)
    T_vec <- numeric(B)
    
    t0 <- proc.time()[3]
    
    #
    for (b in 1:B) {
      eta <- eta0
      V_mat <- matrix(0, 2, 2)   
      
      ## General iteration
      for (t in 1:M) {
        x_new <- rvMF2(eta)
        s <- score_fn(eta, x_new)        
        wt <- 1/(t + n)
        V_mat <- V_mat + wt^2 * (s %*% t(s))
        eta <- eta + wt * s              
      }
      
      ## hybird step
      w_cur <- sum(1 / ((1:M) + n)^2)
      w_rem <- sum(1 / ((M+1):(M+10000) + n)^2)
      Sigma_hat <- V_mat * (1 + w_rem / w_cur)   
      
      eig <- eigen(Sigma_hat, symmetric = TRUE)
      eig$values[eig$values < 0] <- 0
      inv_sqrt_Sigma <- eig$vectors %*% diag(1/sqrt(pmax(eig$values, 1e-12))) %*% t(eig$vectors)
      
      eps <- rnorm(2)
      eta_final <- eta + eig$vectors %*% diag(sqrt(eig$values * w_rem / (w_cur + w_rem))) %*% eps
      
      ## Calculate each stat for UQ.
      delta <- eta_final - eta0
      Z_vec <- inv_sqrt_Sigma %*% delta
      Z_mat[b, ] <- Z_vec
      T_vec[b] <- sum(Z_vec^2)  # For rough test, no rigirous proof.
    }
    t1 <- proc.time()[3]
    times[i] <- t1 - t0
    
    Z_list[[i]] <- Z_mat
    Tn_vec <- c(Tn_vec, T_vec)
  }
  
  Z_all <- do.call(rbind, Z_list)
  colnames(Z_all) <- c("Z1", "Z2")
  
  cover_Z1 <- mean(abs(Z_all[,1]) <= qnorm(0.975))
  cover_Z2 <- mean(abs(Z_all[,2]) <= qnorm(0.975))
  cover_Tn <- mean(Tn_vec <= qchisq(0.95, df = 2))
  
  list(
    Z = Z_all,
    Tn = Tn_vec,
    summary = data.frame(
      Coverage_Z1 = cover_Z1,
      Coverage_Z2 = cover_Z2,
      Coverage_Tn = cover_Tn,
      Mean_Time  = mean(times),
      stringsAsFactors = FALSE
    )
  )
}
```

Note here for simplicity, I only test on with 500 chains

```{r}
res_mart10 <- sim_martingale(n = 10, n_sim = 100, B = 500, M = 50)
res_mart30 <- sim_martingale(n = 30, n_sim = 100, B = 500, M = 50)
res_mart30 <- sim_martingale(n = 100, n_sim = 100, B = 500, M = 50)

result_table <- rbind(res_mart10, res_mart30, res_mart100)
```
