# Simulation Study: Bayesian vMF models for current direction Data with p=2

This vignette demonstrates a complete simulation study using `bayesdir` on synthetic traffic data from a grid road network.

```{r setup, include = FALSE,eval = FALSE,purl=FALSE}
library(bayesdir)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork) 
library(RColorBrewer)
library(rstiefel) 
library(circular) ## for the coordinates in polar system 
library(circglmbayes) ## for sampling kappa through mcmc method. 
library(parallel)
library(coda) ## for effienct sample size

set.seed(20260509)
```

------------------------------------------------------------------------

## Sythetic data for illustration with p=2

Generate data from von-mises fisher distribution. return in both cartesian coordinate and polar coordinate.

Here $\mu \in [0,2\pi]$ as an angle. $\kappa \geq0$ as concentration.

An example,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
data <- generate_vmf_data(100, 2, 4)
```

For p \>2, simply follow $\texttt{rstiefel::rmf.vector}$, below we have an example.

```{r,eval=FALSE,include=FALSE,purl=FALSE}
X <- replicate(n,rmf.vector(c(1,1,1)))
```

## Bayesian Inference on a conjugate prior with p=2.

Used as an benchmark. See Forbes & Mardia (2015) by $\texttt{circglmbayes}$.

For the von Mises distribution on $[0,2\pi)$, $f(\theta\mid\mu,\kappa) = \frac{1}{2\pi I_0(\kappa)}\exp\{\kappa\cos(\theta-\mu)\}$, and conjugate prior $\pi(\mu,\kappa) \propto I_0(\kappa)^{-c}\exp\{\kappa R_0\cos(\mu-\mu_0)\},$ with hyperparameters $c\ge0$, $R_0\ge0$, and $\mu_0$. The resulting joint posterior is$$\pi(\mu,\kappa\mid\theta_{1:n}) \propto I_0(\kappa)^{-(c+n)}\exp{\kappa R_n\cos(\mu-\bar\theta_n)}$$ where the posterior resultant length $R_n$ and mean direction $\bar\theta_n$ are defined by $R_n\cos\bar\theta_n = R_0\cos\mu_0 + \sum_{i=1}^{n}\cos\theta_i$ and $R_n\sin\bar\theta_n = R_0\sin\mu_0 + \sum_{i=1}^{n}\sin\theta_i$.

We sample from the posterior using a two‑step Gibbs sampler.\
Given current values $(\mu,\kappa)$,

-   Draw $\mu\mid\kappa,\theta$ from a von Mises distribution with mean direction $\bar\theta_n$ and concentration $\kappa R_n$.\

-   Draw $\kappa\mid\mu,\theta$ from the full conditional $$    \pi(\kappa\mid\mu,\theta) \propto I_0(\kappa)^{-\eta}\exp\{-\eta\beta_0\kappa\},
      $$ where $\eta = c+n$ and $\beta_0 = -\frac{R_n}{\eta}\cos(\mu-\bar\theta_n)$.

Here $\mu \in [0,2\pi]$ as an angle. $\kappa \geq0$ as concentration. Also, $\texttt{theta}$ as an vector with each scalar $\theta_i\in [0, 2\pi]$ and dimension n.

```{r,eval=FALSE,include=FALSE,purl=FALSE}
fit_mcmc_vmf <- function(theta, prior = list(c = 0, R0 = 0, mu0 = 0), n_iter = 10000, burn = 5000) {
  n <- length(theta)
  C_n <- prior$R0 * cos(prior$mu0) + sum(cos(theta))
  S_n <- prior$R0 * sin(prior$mu0) + sum(sin(theta))
  Rn <- sqrt(C_n^2 + S_n^2)
  theta_bar_n <- atan2(S_n, C_n)
  eta_param <- prior$c + n
  
  mu_sam <- numeric(n_iter); kappa_sam <- numeric(n_iter)
  mu_sam[1] <- theta_bar_n; kappa_sam[1] <- 1
  
  for (iter in 2:n_iter) {
    mu_sam[iter] <- as.numeric(generate_vmf_data(1,theta_bar_n,kappa_sam[iter - 1] * Rn)$theta)
    etag <- -Rn * cos(mu_sam[iter] - theta_bar_n)
    kappa_sam[iter] <- circglmbayes:::sampleKappa(etag = etag, eta = eta_param)[1]
  }
  cbind(mu = mu_sam[-(1:burn)], kappa = kappa_sam[-(1:burn)])
}
```

Below we will show an brief example.

```{r,eval=FALSE,include=FALSE,purl=FALSE}
mu_true <- 2
kappa_true <- 4
data <- generate_vmf_data(100, mu_true = mu_true, kappa_true = kappa_true)
res_MCMC <- fit_mcmc_vmf(data$theta)

kappa_MCMC <- res_MCMC[,2]
mu_MCMC <- res_MCMC[,1]

ess_mu    <- effectiveSize(mu_MCMC)
ess_kappa <- effectiveSize(kappa_MCMC)
cat("Effective Sample Size (after burn-in):\n")
cat(sprintf("  mu    : %.1f\n", ess_mu))
cat(sprintf("  kappa : %.1f\n", ess_kappa))
```

Visualize Result after fitting data,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
plot(kappa_MCMC,type = "l",xlab = "Iteration", ylab = "kappa", main = "Trace of kappa")
abline(h = kappa_true,lty=4,col = "red")

kappa_den <- density(kappa_MCMC)
plot(kappa_den,main = "Posterior density of kappa", xlab = "kappa",
         col = "steelblue", lwd = 2)
abline(v = kappa_true,lty=4,col = "red")


plot(mu_MCMC,type = 'l',xlab = "Iteration", ylab = "mu", main = "Trace of mu")
abline(h = mu_true,lty=4,col = "red")

mu_den <- density(mu_MCMC)
plot(mu_den,main = "Posterior density of mu", xlab = "mu",
         col = "steelblue", lwd = 2)
abline(v = mu_true,lty=4,col = "red")
```

## A simple example for MPS to help you draw graphies in the paper

Though $\texttt{bayesdir::run_predresamp_vmf()}$ in $\texttt{bayesdir()}$

```{r,eval=FALSE,include=FALSE,purl=FALSE}
X <- data$X
nn_cores <- max(detectCores()-1,1)
res_mart <- bayesdir::run_predresamp_vmf(X,B=500,M=1000,n_cores = nn_cores)

kappa_mart <- res_mart$sample$kappa_sam
mu_mart <- res_mart$sample$mu_sam
```

Visualize Result after fitting data,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
kappa_den <- density(kappa_mart)
plot(kappa_den,main = "Posterior density of kappa", xlab = "kappa",
         col = "steelblue", lwd = 2)
abline(v = kappa_true,lty=4,col = "red")


mu_den <- density(mu_mart)
plot(mu_den,main = "Posterior density of mu", xlab = "mu",
         col = "steelblue", lwd = 2)
abline(v = mu_true,lty=4,col = "red")
```

So to draw similar graphes with the paper, follow the comment below carefully

```{r,eval=FALSE,include=FALSE,purl=FALSE}

set.seed(2026) #for n=100,500

#Settings
nn_cores <- max(detectCores()-1,1)
mu_true <- 2
kappa_true <- 4
n <- 500 # use n=10,100,500 
data <- generate_vmf_data(n, mu_true = mu_true, kappa_true = kappa_true)

#MCMC
res_MCMC <- fit_mcmc_vmf(data$theta,n_iter=6000,burn = 5000)
kappa_MCMC <- res_MCMC[,2]
kappa_mcmc_den <- density(kappa_MCMC)
ref_fun <- approxfun(kappa_mcmc_den$x, kappa_mcmc_den$y, yleft = 0, yright = 0)

#MPS
cor_res <- run_predresamp_vmf(data$X, B = 1000, M = 100, type = "HRS-PC-MPS", n_cores = nn_cores, save_raw = TRUE) # 100 for HRS 1000 for RS

kappa_paths <- t(sapply(cor_res$raw_chains, function(chain) {
  sqrt(colSums(chain$c_chain^2))}))

c0 <- bayesdir::mle_vmf(data$X)   
kappa0 <- sqrt(sum(c0^2))
kappa_paths_with_init <- cbind(kappa0, kappa_paths)
kappa_last <- kappa_paths_with_init[, ncol(kappa_paths_with_init)]

#Draw graph, Finally!
layout(matrix(c(1, 2, 3, 3), 2, 2, byrow = TRUE), 
       widths = c(2.4, 1), heights = c(0.9, 0.1))

par(mar = c(3.5, 4, 1.5, 6))

plot_trace_sideways_density(
  paths          = kappa_paths_with_init,
  samples        = kappa_last,          
  iter           = 1:1001, #1:1001 for basic MPS, 1:101 for Hybrid
  ref_value      = kappa_true, 
  main = "Raw Sample Path",
  ylab = "kappa"
)

par(xpd = NA)
usr <- par("usr")
arrow_x0 <- usr[2] + (usr[2] - usr[1]) * 0.08
arrow_x1 <- usr[2] + (usr[2] - usr[1]) * 0.25
arrow_y  <- mean(usr[3:4])

arrows(x0 = arrow_x0, y0 = arrow_y, 
       x1 = arrow_x1, y1 = arrow_y, 
       col = "gray40", lwd = 1.8, length = 0.08)


par(xpd = FALSE)

par(mar = c(4, 1, 1.5, 2)) 
plot_trace_sideways_density(
  paths = NULL,           
  samples = cor_res$sample$kappa_sam,
  ref_density_fun = ref_fun,
  ref_density_col = "red",  
  ref_value      = kappa_true,
  main = "Post-hoc Density",
  ylab = ""
)

par(mar = c(0, 0, 0, 0)) 
plot.new() 

legend("top", 
       legend = c("Initial estimator", "True value", "MPS density", "MCMC benchmark"),
       col = c("red", "red", rgb(0.20, 0.35, 0.65), "red"),
       pch = c(19, NA, 15, NA), 
       lty = c(NA, 2, NA, 1),
       lwd = c(NA, 1.5, NA, 2),
       pt.cex = c(1.2, NA, 1.8, NA),
       bty = "n", 
       horiz = TRUE,   
       cex = 0.9,     
       x.intersp = 0.5,
       text.width = c(0.18,0.12,0.15,0.18)) 

```

## Comparison between MCMC and MPS Method.

As usual, here first give out the function.

```{r,eval=FALSE,include=FALSE,purl=FALSE}
sim_compare_vmf  <- function(n, n_sim = 5000,
                                   mu_true = 2, kappa_true = 4,
                                   B = 5000, M = 1000,
                                   prior = list(c = 0, R0 = 0, mu0 = 0),
                                   seed = 1234, n_cores = 1,
                                   save_dir = "sim_results",
                                   save_interval = n_sim/5) {
  
  if (!dir.exists(save_dir)) {
    dir.create(save_dir, recursive = TRUE)
  }
  
  set.seed(seed)

  
  pb <- txtProgressBar(min = 0, max = n_sim, style = 3)
  start_time <- Sys.time()
  
  # Storage for results
  res_hrs   <- vector("list", n_sim) # Hybrid Resampling (HRS-PC)
  res_rspc  <- vector("list", n_sim) # Standard Resampling (RS-PC)
  res_bayes <- vector("list", n_sim) # MCMC
  
  for (i in 1:n_sim) {
    # 1. Generate Data
    data <- generate_vmf_data(n, mu_true, kappa_true)
    
    # --- A. Martingale: Hybrid (HRS-PC) ---
    t0 <- proc.time()[3]
    hrs_out <- run_predresamp_vmf(data$X, B = B, M = M/10, type = "HRS-PC", cal = "pooled", n_cores = n_cores)
    t1 <- proc.time()[3]
    
    s_mu_hrs <- get_stats(hrs_out$samples$mu_sam, mu_true)
    s_ka_hrs <- get_stats(hrs_out$samples$kappa_sam, kappa_true)
    res_hrs[[i]] <- data.frame(
      c_mu = s_mu_hrs$cover, l_mu = s_mu_hrs$len, v_mu = s_mu_hrs$est_var,
      c_ka = s_ka_hrs$cover, l_ka = s_ka_hrs$len, v_ka = s_ka_hrs$est_var,
      time = t1 - t0
    )

    # --- B. Martingale: Standard (RS-PC) ---
    t2 <- proc.time()[3]
    rs_out <- run_predresamp_vmf(data$X, B = B, M = M, type = "RS-PC", cal = "pooled", n_cores = n_cores)
    t3 <- proc.time()[3]
    
    s_mu_rs <- get_stats(rs_out$samples$mu_sam, mu_true)
    s_ka_rs <- get_stats(rs_out$samples$kappa_sam, kappa_true)
    res_rspc[[i]] <- data.frame(
      c_mu = s_mu_rs$cover, l_mu = s_mu_rs$len, v_mu = s_mu_rs$est_var,
      c_ka = s_ka_rs$cover, l_ka = s_ka_rs$len, v_ka = s_ka_rs$est_var,
      time = t3 - t2
    )

    # --- C. Bayes MCMC ---
    t4 <- proc.time()[3]
    b_samples <- fit_mcmc_vmf(data$theta, prior, n_iter = B + 5000, burn = 5000)
    t5 <- proc.time()[3]
    
    s_mu_b <- get_stats(b_samples[, "mu"], mu_true)
    s_ka_b <- get_stats(b_samples[, "kappa"], kappa_true)
    res_bayes[[i]] <- data.frame(
      c_mu = s_mu_b$cover, l_mu = s_mu_b$len, v_mu = s_mu_b$est_var,
      c_ka = s_ka_b$cover, l_ka = s_ka_b$len, v_ka = s_ka_b$est_var,
      time = t5 - t4
    )
    
    setTxtProgressBar(pb, i)

    # Periodic Backup
    if (i %% save_interval == 0) {
      temp_res <- list(hrs = res_hrs[1:i], rs = res_rspc[1:i], bayes = res_bayes[1:i], iter = i)
      saveRDS(temp_res, file = file.path(save_dir, paste0("backup_iter_", i, ".rds")))
      
      elapsed <- as.numeric(difftime(Sys.time(), start_time, units = "mins"))
      eta <- (elapsed / i) * (n_sim - i)
      message(sprintf("\n[Checkpoint %d] ETA: %.1f mins", i, eta))
    }
  }
  
  close(pb)


  final_summary <- rbind(
    summarize_df(res_hrs, "Martingale-Hybrid"),
    summarize_df(res_rspc, "Martingale-RSPC"),
    summarize_df(res_bayes, "Bayes-MCMC")
  )

  return(final_summary)
}
```

As an simple example,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
getwd()
setwd("~/") ## here set folder to save result, also it will be the place holder to save intermediate result for safety.
nn_cores <- max(detectCores()-1,1)
res <- sim_compare_vmf(n=30, n_sim= 100,B=50,M=100,n_cores = nn_cores)
```

As for code in the paper,

```{r,eval=FALSE,include=FALSE,purl=FALSE}
getwd()
setwd("~/") ## here set folder to save result, also it will be the place holder to save intermediate result for safety.
nn_cores <- max(detectCores()-1,1)
res500 <- sim_compare_vmf(n=500,n_sim=500,B=500,M=1000,n_cores = nn_cores)
save(res500, file = "vmf_simulation_results.RData")

res100 <- sim_compare_vmf(n=100,n_sim=500,B=500,M=1000,n_cores = nn_cores)
save(res500,res100, file = "vmf_simulation_results.RData")

res10 <- sim_compare_vmf(n=10,n_sim=500,B=500,M=1000,n_cores = nn_cores)
save(res500,res100,res10, file = "vmf_simulation_results.RData")
```
