# OSCAR real-data application

This vignette reproduces the real-data results used in the OSCAR southern
offshore compact application: the numerical summary table, the post-corrected
trace-and-terminal-density figure, and the two-panel rose/density figure.  The
analysis uses the vMF predictive-resampling functions in our package; the
small helper routines below only store full trajectories and produce the paper
figures.

```{r,include = FALSE,eval = FALSE,purl=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)

set.seed(2026)
```

```{r packages, include = FALSE,eval = FALSE,purl=FALSE}
library(ggplot2)
library(cowplot)
library(parallel)
library(rstiefel)

load_bayesdir_functions <- function() {
  e <- new.env(parent = emptyenv())

  if (requireNamespace("bayesdir", quietly = TRUE)) {
    e$get_Ap <- bayesdir::get_Ap
    e$score_fn <- bayesdir::score_fn
    e$get_FIM_vmf <- bayesdir::get_FIM_vmf
    e$mle_vmf <- bayesdir::mle_vmf
    e$run_predresamp_vmf <- bayesdir::run_predresamp_vmf
    return(e)
  }

  # Fallback for rendering from the repository root before installing the package.
  source_files <- c("bayesdir/R/all-helpers.R", "bayesdir/R/samp.R")
  if (!all(file.exists(source_files))) {
    stop(
      "Install bayesdir or render this file from the repository root so that ",
      "bayesdir/R/all-helpers.R and bayesdir/R/samp.R are available."
    )
  }
  for (f in source_files) source(f, local = e)
  e
}

.bd <- load_bayesdir_functions()
```

## Data and tuning

The two CSV files contains the preprocessed compact subarea observations
for summer and winter, with columns `y1` and `y2` giving unit current-direction
vectors after rotation to the common reference direction.  For local testing,
the code also searches the experiment folder used to prepare the paper figures.

```{r paths-and-controls, include = FALSE,eval = FALSE,purl=FALSE}
plan_key <- "southern_offshore_compact"

data_dir_candidates <- c(
  file.path("data", "oscar_southern_offshore_compact"),
  file.path("..", "..", "real-data", "spatial_rotation", "reprepared",
            "selected_plan_data", plan_key)
)

data_dir <- data_dir_candidates[file.exists(file.path(data_dir_candidates, "summer_subarea.csv"))][1]
if (is.na(data_dir)) {
  stop("Cannot find summer_subarea.csv. Please place the OSCAR compact subarea data in data/oscar_southern_offshore_compact/.")
}

figure_dir <- file.path("figure", "oscar-real-data")
dir.create(figure_dir, recursive = TRUE, showWarnings = FALSE)

results_rds <- file.path(figure_dir, paste0(plan_key, "_raw_and_postcorrect_results.rds"))
paper_results_rds <- file.path(
  "..", "..", "real-data", "spatial_rotation", "reprepared",
  "southern_offshore_compact_postcorrect_figures",
  paste0(plan_key, "_raw_and_postcorrect_results.rds")
)

quick_mode <- TRUE
nmc <- if (quick_mode) 1200 else 3000
n_chains <- if (quick_mode) 800 else 5000
trace_show <- if (quick_mode) 150 else 250
detected_cores <- parallel::detectCores()
n_cores <- if (is.na(detected_cores)) 1 else max(1, detected_cores - 1)

# Set to TRUE to rerun predictive resampling even if a cached result file exists.
rerun_mps <- FALSE

rose_bins <- 36
rose_inner_radius <- 0.28
```

## Helper functions

```{r helpers, include = FALSE,eval = FALSE,purl=FALSE}
unit_matrix <- function(df) {
  t(as.matrix(df[, c("y1", "y2")]))
}

obs_matrix <- function(X) {
  t(X)
}

angle_to_signed_degrees <- function(theta) {
  out <- theta * 180 / pi
  ((out + 180) %% 360) - 180
}

fit_vmf <- function(X) {
  c_hat <- as.numeric(.bd$mle_vmf(X))
  kappa_hat <- sqrt(sum(c_hat^2))
  mu_hat <- if (kappa_hat > 0) c_hat / kappa_hat else rep(0, length(c_hat))
  x_bar <- rowMeans(X)
  R_bar <- sqrt(sum(x_bar^2))

  list(
    mu_hat = mu_hat,
    kappa_hat = kappa_hat,
    c_hat = c_hat,
    R_bar = R_bar,
    x_bar = x_bar
  )
}

extract_posterior <- function(c_array) {
  n_iter <- dim(c_array)[1]
  c_terminal <- t(c_array[n_iter, , ])
  kappa <- sqrt(rowSums(c_terminal^2))
  phi <- atan2(c_terminal[, 2], c_terminal[, 1])
  data.frame(
    c1 = c_terminal[, 1],
    c2 = c_terminal[, 2],
    kappa = kappa,
    phi_rad = phi,
    phi_deg = angle_to_signed_degrees(phi)
  )
}

make_summary <- function(post, mle, N, season, posterior_label) {
  data.frame(
    Posterior = posterior_label,
    Season = season,
    N = N,
    R_bar = round(mle$R_bar, 4),
    kappa_MLE = round(mle$kappa_hat, 4),
    kappa_mean = round(mean(post$kappa), 4),
    kappa_sd = round(sd(post$kappa), 4),
    kappa_q025 = round(quantile(post$kappa, 0.025), 4),
    kappa_q975 = round(quantile(post$kappa, 0.975), 4),
    phi_MLE_deg = round(angle_to_signed_degrees(atan2(mle$c_hat[2], mle$c_hat[1])), 2),
    phi_mean_deg = round(mean(post$phi_deg), 2),
    phi_sd_deg = round(sd(post$phi_deg), 2),
    row.names = NULL
  )
}

safe_sym_inverse <- function(M, tol = 1e-10) {
  eig <- eigen((M + t(M)) / 2, symmetric = TRUE)
  vals_inv <- 1 / pmax(eig$values, tol)
  eig$vectors %*% diag(vals_inv, nrow = length(vals_inv)) %*% t(eig$vectors)
}

parallel_lapply <- function(X, FUN, ..., n_cores = 1) {
  if (.Platform$OS.type == "windows" || n_cores <= 1) {
    lapply(X, FUN, ...)
  } else {
    parallel::mclapply(X, FUN, ..., mc.cores = n_cores, mc.preschedule = TRUE)
  }
}

run_single_path <- function(seed, c0, N, M) {
  set.seed(seed)
  p <- length(c0)
  path <- matrix(0, nrow = M + 1, ncol = p)
  path[1, ] <- c0

  for (t in seq_len(M)) {
    c_curr <- path[t, ]
    x_new <- as.numeric(rstiefel::rmf.vector(c_curr))
    score <- .bd$score_fn(c_curr, x_new, p)
    wt <- 1 / (t + N - 1)
    path[t + 1, ] <- c_curr + wt * score
  }

  path
}

run_mps_paths <- function(X, M, B, n_cores = 1, base_seed = 42) {
  c0 <- as.numeric(.bd$mle_vmf(X))
  paths <- parallel_lapply(
    seq_len(B),
    function(i) run_single_path(base_seed + i, c0 = c0, N = ncol(X), M = M),
    n_cores = n_cores
  )

  out <- array(0, dim = c(M + 1, nrow(X), B))
  for (i in seq_len(B)) out[, , i] <- paths[[i]]
  out
}

post_correct_single_path <- function(path, N, p = ncol(path), tol = 1e-10,
                                     correct_full_path = TRUE) {
  c0 <- path[1, ]
  V <- matrix(0, p, p)
  total_weight <- 0

  for (m_idx in seq_len(nrow(path) - 1)) {
    m <- N + m_idx - 1
    wt2 <- (1 / m)^2
    V <- V + wt2 * .bd$get_FIM_vmf(path[m_idx, ], p)
    total_weight <- total_weight + wt2
  }

  A <- total_weight * safe_sym_inverse(V, tol = tol)
  if (!correct_full_path) {
    out <- path
    out[nrow(path), ] <- as.vector(c0 + A %*% (path[nrow(path), ] - c0))
    return(out)
  }

  centered <- t(t(path) - c0)
  corrected <- t(A %*% t(centered))
  t(t(corrected) + c0)
}

post_correct_array <- function(c_array, N, p = dim(c_array)[2], n_cores = 1) {
  corrected <- parallel_lapply(
    seq_len(dim(c_array)[3]),
    function(i) post_correct_single_path(c_array[, , i], N = N, p = p),
    n_cores = n_cores
  )

  out <- array(0, dim = dim(c_array))
  for (i in seq_along(corrected)) out[, , i] <- corrected[[i]]
  out
}
```

## Load data and compute MLEs

```{r load-data, include = FALSE,eval = FALSE,purl=FALSE}
summer_df <- read.csv(file.path(data_dir, "summer_subarea.csv"))
winter_df <- read.csv(file.path(data_dir, "winter_subarea.csv"))

X_summer <- unit_matrix(summer_df)
X_winter <- unit_matrix(winter_df)
X_summer_obs <- obs_matrix(X_summer)
X_winter_obs <- obs_matrix(X_winter)

mle_summer <- fit_vmf(X_summer)
mle_winter <- fit_vmf(X_winter)

mle_table <- data.frame(
  Season = c("Summer", "Winter"),
  N = c(ncol(X_summer), ncol(X_winter)),
  R_bar = c(mle_summer$R_bar, mle_winter$R_bar),
  mu1_hat = c(mle_summer$mu_hat[1], mle_winter$mu_hat[1]),
  mu2_hat = c(mle_summer$mu_hat[2], mle_winter$mu_hat[2]),
  phi_hat_deg = c(
    angle_to_signed_degrees(atan2(mle_summer$mu_hat[2], mle_summer$mu_hat[1])),
    angle_to_signed_degrees(atan2(mle_winter$mu_hat[2], mle_winter$mu_hat[1]))
  ),
  kappa_hat = c(mle_summer$kappa_hat, mle_winter$kappa_hat)
)

mle_table_print <- mle_table
num_cols <- vapply(mle_table_print, is.numeric, logical(1))
mle_table_print[num_cols] <- lapply(mle_table_print[num_cols], round, 4)
knitr::kable(mle_table_print, caption = "Seasonal vMF maximum likelihood estimates.")
```

## Predictive resampling and post-correction

The chunk below reruns the full predictive-resampling experiment when
`rerun_mps = TRUE` or when no cached result file is available.  The full-path
helper uses the same score, Fisher information, and affine variance correction
as the exported `bayesdir` sampler, but stores the intermediate states needed
for the trace figure.

```{r run-or-load-results, include = FALSE,eval = FALSE,purl=FALSE}
cached_source <- if (file.exists(results_rds)) results_rds else paper_results_rds

if (!rerun_mps && file.exists(cached_source)) {
  res <- readRDS(cached_source)
} else {
  arr_summer_raw <- run_mps_paths(X_summer, M = nmc, B = n_chains, n_cores = n_cores)
  arr_winter_raw <- run_mps_paths(X_winter, M = nmc, B = n_chains, n_cores = n_cores)

  arr_summer_pc <- post_correct_array(arr_summer_raw, N = ncol(X_summer), n_cores = n_cores)
  arr_winter_pc <- post_correct_array(arr_winter_raw, N = ncol(X_winter), n_cores = n_cores)

  res <- list(
    mle_summer = mle_summer,
    mle_winter = mle_winter,
    post_summer_raw = extract_posterior(arr_summer_raw),
    post_winter_raw = extract_posterior(arr_winter_raw),
    post_summer_pc = extract_posterior(arr_summer_pc),
    post_winter_pc = extract_posterior(arr_winter_pc),
    arr_summer_raw = arr_summer_raw,
    arr_winter_raw = arr_winter_raw,
    arr_summer_pc = arr_summer_pc,
    arr_winter_pc = arr_winter_pc
  )

  saveRDS(res, results_rds)
}

mle_summer <- res$mle_summer
mle_winter <- res$mle_winter
post_summer_raw <- res$post_summer_raw
post_winter_raw <- res$post_winter_raw
post_summer_pc <- res$post_summer_pc
post_winter_pc <- res$post_winter_pc
arr_summer_pc <- res$arr_summer_pc
arr_winter_pc <- res$arr_winter_pc

summary_table <- rbind(
  make_summary(res$post_summer_raw, mle_summer, ncol(X_summer), "Summer", "Raw MPS"),
  make_summary(res$post_winter_raw, mle_winter, ncol(X_winter), "Winter", "Raw MPS"),
  make_summary(post_summer_pc, mle_summer, ncol(X_summer), "Summer", "Post-corrected MPS"),
  make_summary(post_winter_pc, mle_winter, ncol(X_winter), "Winter", "Post-corrected MPS")
)

write.csv(summary_table, file.path(figure_dir, "oscar_southern_offshore_compact_summary_table.csv"), row.names = FALSE)
knitr::kable(summary_table, caption = "Raw and post-corrected martingale posterior summaries.")
```

## Post-corrected trace plot

```{r trace-plot, fig.width=15, fig.height=5.625, include = FALSE,eval = FALSE,purl=FALSE}
trace_axis_label_cex <- 1.8
trace_axis_tick_cex <- 1.6
trace_legend_location_units <- "npc"
trace_legend_location_kappa <- c(-0.18, 1.00)
trace_legend_location_phi <- c(-0.18, 1.00)
trace_legend_cex <- 1.25
trace_legend_xpd <- NA
trace_subplot_mar <- c(4.2, 5.8, 2.4, 1.4)
trace_outer_mar <- c(0.5, 0.5, 2.0, 0)
trace_panel_widths <- c(1.00, 1.00)
trace_panel_respect <- FALSE
trace_subtitle_cex <- 1.8
trace_outer_title_cex <- 1.10
trace_line_lwd <- 0.9
trace_mle_lwd <- 1.5
trace_density_lwd <- 1.3
trace_xlim_multiplier <- 1.38
trace_density_x_offset <- 1.04
trace_density_width <- 0.28
kappa_ylim <- c(1.1, 3.25)
phi_use_compressed_y_axis <- TRUE
phi_lower_ylim <- NULL
phi_upper_ylim <- c(90, 160)
phi_gap_fraction <- 0.16
phi_gap_label_cex <- 0.70

trace_matrices <- function(c_array, n_show) {
  n_show <- min(n_show, dim(c_array)[3])
  list(
    kappa = apply(c_array[, , seq_len(n_show), drop = FALSE], c(1, 3), function(v) sqrt(sum(v^2))),
    phi = apply(c_array[, , seq_len(n_show), drop = FALSE], c(1, 3), function(v) angle_to_signed_degrees(atan2(v[2], v[1])))
  )
}

pad_range <- function(x, frac = 0.08, min_pad = 0.1) {
  r <- range(x, finite = TRUE)
  d <- diff(r)
  pad <- max(frac * d, min_pad)
  c(r[1] - pad, r[2] + pad)
}

make_compressed_axis <- function(arr_summer, arr_winter, lower_ylim = NULL,
                                 upper_ylim = NULL, gap_fraction = 0.16) {
  r_s <- pad_range(arr_summer)
  r_w <- pad_range(arr_winter)
  if (mean(r_s) <= mean(r_w)) {
    lower <- if (is.null(lower_ylim)) r_s else lower_ylim
    upper <- if (is.null(upper_ylim)) r_w else upper_ylim
  } else {
    lower <- if (is.null(lower_ylim)) r_w else lower_ylim
    upper <- if (is.null(upper_ylim)) r_s else upper_ylim
  }

  lower <- sort(lower)
  upper <- sort(upper)
  lower_height <- diff(lower)
  upper_height <- diff(upper)
  gap_height <- max(lower_height, upper_height, 1e-8) * gap_fraction

  map_y <- function(y) {
    y_mapped <- numeric(length(y))
    lower_idx <- y <= lower[2]
    upper_idx <- y >= upper[1]
    gap_idx <- !(lower_idx | upper_idx)
    y_mapped[lower_idx] <- y[lower_idx] - lower[1]
    y_mapped[upper_idx] <- lower_height + gap_height + (y[upper_idx] - upper[1])
    if (any(gap_idx)) {
      y_mapped[gap_idx] <- lower_height +
        (y[gap_idx] - lower[2]) / max(upper[1] - lower[2], 1e-8) * gap_height
    }
    y_mapped
  }

  ticks <- c(pretty(lower), pretty(upper))
  ticks <- ticks[(ticks >= lower[1] & ticks <= lower[2]) | (ticks >= upper[1] & ticks <= upper[2])]

  list(
    lower_height = lower_height,
    gap_height = gap_height,
    y_limits_mapped = c(0, lower_height + gap_height + upper_height),
    map_y = map_y,
    ticks = ticks,
    tick_positions = map_y(ticks)
  )
}

draw_trace_x_axis <- function(n_iter, label = "Forward step", line = 2.8) {
  ticks <- pretty(c(0, n_iter))
  ticks <- ticks[ticks >= 0 & ticks <= n_iter]
  axis(1, at = ticks, labels = ticks, lwd = 0, lwd.ticks = 1, cex.axis = trace_axis_tick_cex)
  usr <- par("usr")
  segments(0, usr[3], n_iter, usr[3], xpd = FALSE)
  mtext(label, side = 1, line = line, at = n_iter / 2, cex = trace_axis_label_cex)
}

legend_location_to_plot <- function(location, map_y = identity) {
  if (!is.numeric(location) || length(location) != 2 || trace_legend_location_units != "npc") {
    return(location)
  }
  usr <- par("usr")
  c(usr[1] + location[1] * diff(usr[1:2]), usr[3] + location[2] * diff(usr[3:4]))
}

combined_trace_panel <- function(arr_summer, arr_winter, mle_summer_val, mle_winter_val,
                                 ylab_expr, label, col_s, fill_s, border_s,
                                 col_w, fill_w, border_w, legend_location,
                                 y_limits = NULL, compressed_axis = NULL) {
  n_iter <- nrow(arr_summer)
  iters <- seq_len(n_iter)
  x_limits <- c(0, n_iter * trace_xlim_multiplier)

  if (is.null(compressed_axis)) {
    map_y <- function(y) y
    y_plot_limits <- if (is.null(y_limits)) pad_range(c(arr_summer, arr_winter)) else y_limits
    y_axis_ticks <- pretty(y_plot_limits)
    y_axis_positions <- y_axis_ticks
  } else {
    map_y <- compressed_axis$map_y
    y_plot_limits <- compressed_axis$y_limits_mapped
    y_axis_ticks <- compressed_axis$ticks
    y_axis_positions <- compressed_axis$tick_positions
  }

  plot(NULL, xlim = x_limits, ylim = y_plot_limits,
       xlab = "", ylab = ylab_expr, main = label,
       xaxt = "n", yaxt = "n",
       cex.lab = trace_axis_label_cex, cex.main = trace_subtitle_cex)
  draw_trace_x_axis(n_iter)
  axis(2, at = y_axis_positions, labels = y_axis_ticks, cex.axis = trace_axis_tick_cex)

  for (i in seq_len(ncol(arr_summer))) lines(iters, map_y(arr_summer[, i]), col = col_s, lwd = trace_line_lwd)
  for (i in seq_len(ncol(arr_winter))) lines(iters, map_y(arr_winter[, i]), col = col_w, lwd = trace_line_lwd)
  abline(h = map_y(mle_summer_val), lty = 2, col = border_s, lwd = trace_mle_lwd)
  abline(h = map_y(mle_winter_val), lty = 2, col = border_w, lwd = trace_mle_lwd)

  dens_s <- density(arr_summer[n_iter, ], n = 512)
  dens_w <- density(arr_winter[n_iter, ], n = 512)
  dens_scale <- n_iter * trace_density_width / max(c(dens_s$y, dens_w$y))
  x_offset <- n_iter * trace_density_x_offset
  polygon(x_offset + dens_s$y * dens_scale, map_y(dens_s$x), col = fill_s, border = border_s, lwd = trace_density_lwd)
  polygon(x_offset + dens_w$y * dens_scale, map_y(dens_w$x), col = fill_w, border = border_w, lwd = trace_density_lwd)
  segments(x_offset, y_plot_limits[1], x_offset, y_plot_limits[2], col = "gray60", lty = 3, lwd = 0.8)

  if (!is.null(compressed_axis)) {
    gap_mid <- compressed_axis$lower_height + compressed_axis$gap_height / 2
    abline(h = gap_mid, col = "gray78", lty = 3, lwd = 0.8)
    text(n_iter * 0.52, gap_mid, "compressed y gap", col = "gray45", cex = phi_gap_label_cex)
  }

  legend(legend_location_to_plot(legend_location, map_y),
         legend = c("Summer traces", "Winter traces", "Summer terminal density", "Winter terminal density"),
         col = c(border_s, border_w, border_s, border_w),
         lty = 1, lwd = c(trace_line_lwd, trace_line_lwd, 4, 4),
         bty = "n", cex = trace_legend_cex, xpd = trace_legend_xpd)
}

plot_combined_trace_figure <- function(arr_summer, arr_winter, mle_summer, mle_winter) {
  tm_s <- trace_matrices(arr_summer, trace_show)
  tm_w <- trace_matrices(arr_winter, trace_show)
  phi_axis <- if (isTRUE(phi_use_compressed_y_axis)) {
    make_compressed_axis(tm_s$phi, tm_w$phi, phi_lower_ylim, phi_upper_ylim, phi_gap_fraction)
  } else {
    NULL
  }

  old_par <- par(no.readonly = TRUE)
  on.exit(par(old_par), add = TRUE)
  layout(matrix(c(1, 2), nrow = 1), widths = trace_panel_widths, respect = trace_panel_respect)
  par(mar = trace_subplot_mar, oma = trace_outer_mar)

  combined_trace_panel(
    tm_s$kappa, tm_w$kappa, mle_summer$kappa_hat, mle_winter$kappa_hat,
    expression(kappa), "(a) " ~ kappa,
    rgb(0.25, 0.45, 0.75, 0.06), rgb(0.25, 0.45, 0.75, 0.24), rgb(0.2, 0.35, 0.65),
    rgb(0.9, 0.45, 0.2, 0.06), rgb(0.9, 0.45, 0.2, 0.24), rgb(0.8, 0.35, 0.15),
    trace_legend_location_kappa,
    y_limits = kappa_ylim
  )

  combined_trace_panel(
    tm_s$phi, tm_w$phi,
    angle_to_signed_degrees(atan2(mle_summer$mu_hat[2], mle_summer$mu_hat[1])),
    angle_to_signed_degrees(atan2(mle_winter$mu_hat[2], mle_winter$mu_hat[1])),
    expression(phi ~ "(deg)"), "(b) " ~ phi,
    rgb(0.25, 0.45, 0.75, 0.06), rgb(0.25, 0.45, 0.75, 0.24), rgb(0.2, 0.35, 0.65),
    rgb(0.9, 0.45, 0.2, 0.06), rgb(0.9, 0.45, 0.2, 0.24), rgb(0.8, 0.35, 0.15),
    trace_legend_location_phi,
    compressed_axis = phi_axis
  )
}

trace_png <- file.path(figure_dir, "oscar_southern_offshore_compact_combined_traces_postcorrected.png")
trace_pdf <- file.path(figure_dir, "oscar_southern_offshore_compact_combined_traces_postcorrected.pdf")

png(trace_png, width = 2400, height = 900, res = 160)
plot_combined_trace_figure(arr_summer_pc, arr_winter_pc, mle_summer, mle_winter)
dev.off()

pdf(trace_pdf, width = 15.0, height = 5.625)
plot_combined_trace_figure(arr_summer_pc, arr_winter_pc, mle_summer, mle_winter)
dev.off()

plot_combined_trace_figure(arr_summer_pc, arr_winter_pc, mle_summer, mle_winter)
```

## Rose diagram and angular density

```{r rose-density-figure, fig.width=13.5, fig.height=6.2, include = FALSE,eval = FALSE,purl=FALSE}
paper2_season_colors <- c("Summer" = "#2b6cb0", "Winter" = "#dd6b20")
paper2_hist_alpha <- 0.4
paper2_band_alpha <- 0.16
paper2_hist_edge_alpha <- 0.55
paper2_density_linewidth <- 1.05
paper2_hist_linewidth <- 0.25
paper2_mean_arrow_linewidth <- 0.85
paper2_subtitle_size <- 18
paper2_legend_font_size <- 16
paper2_axis_label_size <- 14
paper2_axis_tick_size <- 13
paper2_polar_tick_size <- 5.0
paper2_gap_pt <- 16
paper2_rel_widths <- c(1.0, 1.0)

circle_xy <- function(r = 1, n = 400) {
  theta <- seq(0, 2 * pi, length.out = n)
  data.frame(x = r * cos(theta), y = r * sin(theta))
}

vmf_density_s1 <- function(theta, kappa, phi) {
  log_dens <- kappa * cos(theta - phi) - log(2 * pi) -
    (log(besselI(kappa, 0, expon.scaled = TRUE)) + kappa)
  exp(log_dens)
}

posterior_density_band <- function(theta_grid, post, max_draws = 600) {
  n_draws <- min(max_draws, nrow(post))
  idx <- if (n_draws < nrow(post)) sample(seq_len(nrow(post)), n_draws) else seq_len(nrow(post))
  dens_mat <- vapply(idx, function(i) vmf_density_s1(theta_grid, post$kappa[i], post$phi_rad[i]), numeric(length(theta_grid)))
  if (is.vector(dens_mat)) dens_mat <- matrix(dens_mat, ncol = 1)
  list(
    mean = rowMeans(dens_mat),
    lo = apply(dens_mat, 1, quantile, probs = 0.025),
    hi = apply(dens_mat, 1, quantile, probs = 0.975)
  )
}

make_rose_hist_polygons <- function(X_mat, season_label, scale_r, n_bins = rose_bins) {
  base_r <- rose_inner_radius
  angles_deg <- atan2(X_mat[, 2], X_mat[, 1]) * 180 / pi
  N <- nrow(X_mat)
  bin_width <- 360 / n_bins
  breaks <- seq(-180, 180, by = bin_width)
  angle_cut <- cut(angles_deg, breaks = breaks, include.lowest = TRUE, right = FALSE)
  counts <- as.numeric(table(angle_cut))
  dens_hist <- counts / (N * bin_width * pi / 180)

  poly_list <- lapply(seq_along(counts), function(i) {
    a_s <- breaks[i] * pi / 180
    a_e <- breaks[i + 1] * pi / 180
    a_seq <- seq(a_s, a_e, length.out = 20)
    r_val <- base_r + scale_r * dens_hist[i]
    data.frame(
      x = c(base_r * cos(a_seq[1]), r_val * cos(a_seq), base_r * cos(rev(a_seq)[1])),
      y = c(base_r * sin(a_seq[1]), r_val * sin(a_seq), base_r * sin(rev(a_seq)[1])),
      bin = paste(season_label, i, sep = "_"),
      season = season_label
    )
  })
  do.call(rbind, poly_list)
}

make_posterior_band_polygon <- function(theta_grid, post_band, season_label, scale_r) {
  base_r <- rose_inner_radius
  r_lo <- base_r + scale_r * post_band$lo
  r_hi <- base_r + scale_r * post_band$hi
  data.frame(
    x = c(r_lo * cos(theta_grid), rev(r_hi * cos(theta_grid))),
    y = c(r_lo * sin(theta_grid), rev(r_hi * sin(theta_grid))),
    season = season_label
  )
}

make_density_curve <- function(theta_grid, mle, season_label, scale_r) {
  phi <- atan2(mle$mu_hat[2], mle$mu_hat[1])
  dens <- vmf_density_s1(theta_grid, mle$kappa_hat, phi)
  r <- rose_inner_radius + scale_r * dens
  data.frame(x = r * cos(theta_grid), y = r * sin(theta_grid), season = season_label)
}

make_mean_arrow <- function(mle, season_label, max_density, scale_r) {
  phi <- atan2(mle$mu_hat[2], mle$mu_hat[1])
  arrow_r <- rose_inner_radius + scale_r * max_density * 0.86
  data.frame(x = 0, y = 0, xend = arrow_r * cos(phi), yend = arrow_r * sin(phi), season = season_label)
}

build_combined_polar_paper <- function(X_summer, X_winter, mle_summer, mle_winter,
                                       post_summer, post_winter, scale_r, tick_r) {
  theta_fine <- seq(-pi, pi, length.out = 500)
  phi_s <- atan2(mle_summer$mu_hat[2], mle_summer$mu_hat[1])
  phi_w <- atan2(mle_winter$mu_hat[2], mle_winter$mu_hat[1])
  dens_s <- vmf_density_s1(theta_fine, mle_summer$kappa_hat, phi_s)
  dens_w <- vmf_density_s1(theta_fine, mle_winter$kappa_hat, phi_w)
  post_band_s <- posterior_density_band(theta_fine, post_summer)
  post_band_w <- posterior_density_band(theta_fine, post_winter)

  hist_df <- rbind(
    make_rose_hist_polygons(X_summer, "Summer", scale_r),
    make_rose_hist_polygons(X_winter, "Winter", scale_r)
  )
  band_df <- rbind(
    make_posterior_band_polygon(theta_fine, post_band_s, "Summer", scale_r),
    make_posterior_band_polygon(theta_fine, post_band_w, "Winter", scale_r)
  )
  curve_df <- rbind(
    make_density_curve(theta_fine, mle_summer, "Summer", scale_r),
    make_density_curve(theta_fine, mle_winter, "Winter", scale_r)
  )
  arrow_df <- rbind(
    make_mean_arrow(mle_summer, "Summer", max(dens_s), scale_r),
    make_mean_arrow(mle_winter, "Winter", max(dens_w), scale_r)
  )

  tick_angles <- c(0, 90, 180, -90) * pi / 180
  tick_df <- data.frame(
    x = tick_r * cos(tick_angles),
    y = tick_r * sin(tick_angles),
    lab = c("0*degree", "90*degree", "180*degree", "-90*degree")
  )

  ggplot() +
    geom_polygon(data = hist_df, aes(x, y, group = bin, fill = season),
                 color = scales::alpha("white", paper2_hist_edge_alpha),
                 linewidth = paper2_hist_linewidth, alpha = paper2_hist_alpha) +
    geom_polygon(data = band_df, aes(x, y, group = season, fill = season),
                 alpha = paper2_band_alpha, color = NA, show.legend = FALSE) +
    geom_path(data = curve_df, aes(x, y, color = season),
              linewidth = paper2_density_linewidth, show.legend = TRUE) +
    geom_path(data = circle_xy(rose_inner_radius), aes(x, y), color = "grey50", linewidth = 0.35) +
    geom_segment(data = arrow_df, aes(x = x, y = y, xend = xend, yend = yend, color = season),
                 arrow = arrow(length = grid::unit(0.15, "cm")),
                 linewidth = paper2_mean_arrow_linewidth, show.legend = FALSE) +
    geom_text(data = tick_df, aes(x, y, label = lab), size = paper2_polar_tick_size, color = "grey30", parse = TRUE) +
    scale_fill_manual(values = paper2_season_colors, name = "Season") +
    scale_color_manual(values = paper2_season_colors, name = "Season") +
    guides(fill = guide_legend(override.aes = list(alpha = 0.55, linewidth = 0)), color = "none") +
    coord_fixed() +
    labs(title = "(a)") +
    theme_void(base_size = paper2_axis_label_size) +
    theme(
      plot.title = element_text(size = paper2_subtitle_size, face = "bold", hjust = 0.5),
      legend.position = "bottom",
      legend.title = element_text(size = paper2_legend_font_size),
      legend.text = element_text(size = paper2_legend_font_size),
      plot.margin = margin(5.5, paper2_gap_pt / 2, 5.5, 5.5)
    )
}

build_angular_density_plot <- function(X_summer, X_winter, mle_summer, mle_winter,
                                       post_summer, post_winter, n_bins = rose_bins) {
  theta_grid <- seq(-pi, pi, length.out = 500)
  obs_df <- rbind(
    data.frame(angle = atan2(X_summer[, 2], X_summer[, 1]) * 180 / pi, season = "Summer"),
    data.frame(angle = atan2(X_winter[, 2], X_winter[, 1]) * 180 / pi, season = "Winter")
  )

  make_density_df <- function(mle, post, season_label) {
    phi <- atan2(mle$mu_hat[2], mle$mu_hat[1])
    dens <- vmf_density_s1(theta_grid, mle$kappa_hat, phi) * (pi / 180)
    post_band <- posterior_density_band(theta_grid, post)
    data.frame(
      angle = theta_grid * 180 / pi,
      density = dens,
      density_lo = post_band$lo * (pi / 180),
      density_hi = post_band$hi * (pi / 180),
      season = season_label
    )
  }

  fit_df <- rbind(
    make_density_df(mle_summer, post_summer, "Summer"),
    make_density_df(mle_winter, post_winter, "Winter")
  )

  ggplot() +
    geom_histogram(data = obs_df, aes(x = angle, y = after_stat(density), fill = season),
                   binwidth = 360 / n_bins, boundary = -180, closed = "left",
                   position = "identity", alpha = paper2_hist_alpha,
                   color = "white", linewidth = 0.25) +
    geom_ribbon(data = fit_df, aes(x = angle, ymin = density_lo, ymax = density_hi, fill = season),
                alpha = paper2_band_alpha) +
    geom_line(data = fit_df, aes(x = angle, y = density, color = season),
              linewidth = paper2_density_linewidth) +
    scale_fill_manual(values = paper2_season_colors, name = "Season") +
    scale_color_manual(values = paper2_season_colors, name = "Season") +
    scale_x_continuous(breaks = seq(-180, 180, 60), limits = c(-180, 180)) +
    labs(title = "(b)", x = expression(Delta * theta ~ "(degrees)"), y = "Density") +
    theme_minimal(base_size = paper2_axis_label_size) +
    theme(
      plot.title = element_text(size = paper2_subtitle_size, face = "bold", hjust = 0.5),
      axis.title = element_text(size = paper2_axis_label_size),
      axis.text = element_text(size = paper2_axis_tick_size),
      legend.position = "bottom",
      legend.title = element_text(size = paper2_legend_font_size),
      legend.text = element_text(size = paper2_legend_font_size),
      plot.margin = margin(5.5, 5.5, 5.5, paper2_gap_pt / 2)
    )
}

plot_paper_figure <- function(X_summer, X_winter, mle_summer, mle_winter,
                              post_summer, post_winter) {
  theta_fine <- seq(-pi, pi, length.out = 500)
  hist_max_fn <- function(X_mat) {
    angles_deg <- atan2(X_mat[, 2], X_mat[, 1]) * 180 / pi
    bin_width <- 360 / rose_bins
    breaks <- seq(-180, 180, by = bin_width)
    angle_cut <- cut(angles_deg, breaks = breaks, include.lowest = TRUE, right = FALSE)
    counts <- as.numeric(table(angle_cut))
    max(counts / (nrow(X_mat) * bin_width * pi / 180))
  }
  phi_s <- atan2(mle_summer$mu_hat[2], mle_summer$mu_hat[1])
  phi_w <- atan2(mle_winter$mu_hat[2], mle_winter$mu_hat[1])
  post_band_s <- posterior_density_band(theta_fine, post_summer)
  post_band_w <- posterior_density_band(theta_fine, post_winter)
  global_max <- max(
    hist_max_fn(X_summer),
    hist_max_fn(X_winter),
    max(vmf_density_s1(theta_fine, mle_summer$kappa_hat, phi_s)),
    max(vmf_density_s1(theta_fine, mle_winter$kappa_hat, phi_w)),
    max(post_band_s$hi),
    max(post_band_w$hi)
  )
  scale_r <- 0.7 / global_max
  tick_r <- rose_inner_radius + scale_r * global_max * 1.12

  p_a <- build_combined_polar_paper(X_summer, X_winter, mle_summer, mle_winter,
                                    post_summer, post_winter, scale_r, tick_r)
  p_b <- build_angular_density_plot(X_summer, X_winter, mle_summer, mle_winter,
                                    post_summer, post_winter)

  plot_grid(p_a, p_b, nrow = 1, rel_widths = paper2_rel_widths, align = "h") +
    theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5))
}

paper_fig <- plot_paper_figure(
  X_summer_obs, X_winter_obs,
  mle_summer, mle_winter,
  post_summer_pc, post_winter_pc
)

rose_png <- file.path(figure_dir, "oscar_southern_offshore_compact_paper_rose_density_postcorrected.png")
rose_pdf <- file.path(figure_dir, "oscar_southern_offshore_compact_paper_rose_density_postcorrected.pdf")

ggsave(rose_png, paper_fig, width = 13.5, height = 6.2, dpi = 220)
ggsave(rose_pdf, paper_fig, width = 13.5, height = 6.2)

paper_fig
```
